On The Subspace Learning for Network Attack Detection

### 2019-06-09 docs/thesis/chapters/1_introduction/introduction.tex
results of robpca

Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}. In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices. Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.

Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data. Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}. Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}. Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.

Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis. However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}. To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.

>>>>

Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}. In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have been applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices. Intrusion detection and intrusion prevention systems are two types of security systems that are being used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.

Anomaly detection can be defined as the identification of rare and suspicious events by comparing it to the normal data. Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}. Additionally, anomaly detection techniques can be categorized by classification, statistical algorithms, information theory and cluster based algorithms, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}. Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events. Therefore anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem \cite{he2008learning}.

Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis. However, distributed attacks organized by botnet has increased and demanded the development of counter measures. This has been done to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns. This has also been done in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}. To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.

### 2019-06-09 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
results of robpca

According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time. The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.

>>>>

According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time. The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7\% and 199\%.
